Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-11-06 manuscript/manuscript.Rmd
Initial implementation of sensitivity-to-outlier plots.

```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=2.0, warning=FALSE}
cap <- paste0(
  "\\label{fig:decreased-normality}",
  "Power transformations may reduce normality depending on location. ",
  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
  "and even breaks down due to numerical issues for $\\mu > 5 \\cdot 10^4$.")

@@

Real-world data may contain outliers, to which maximum likelihood estimation can be sensitive. Their presence may lead to poor transformations to normality, as shown in Figure **INSERT**. As indicated by @Raymaekers2021-kq, the general aim of power transformations should be to transform non-outlier data to normality, i.e. achieve *central normality*. To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$. Because $\lambda$ is updated during their procedure, simultaneous estimation of $\lambda$ and $x_0$ for shift-sensitive power transformations is not possible.

>>>>

```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=2.0, warning=FALSE}
cap <- paste0(
  "\\label{fig:decreased-normality}",
  "Power transformations may reduce normality depending on location. ",
  "Variable $x$ consists of $10000$ values randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
  "and even breaks down due to numerical issues for $\\mu > 5 \\cdot 10^4$.")

@@

```{r sensitivity-to-outliers, echo=FALSE, fig.cap=cap, fig.height=2.0, warning=FALSE}
cap <- paste0(
  "\\label{fig:sensitivity-to-outliers}",
  "Power transformations are affected by outliers. ",
  "Variable $x$ consists of $1000$ values randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$, ",
  "with a single outlier with value $d$. ",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly affected by the presence of the outlier at $d$, ",
  "and even breaks down due to numerical issues for $d > 5 \\cdot 10^4$.")

.plot_sensitivity_to_outlier(plot_theme = plot_theme, manuscript_dir = manuscript_dir)

Real-world data may contain outliers, to which maximum likelihood estimation can be sensitive. Their presence may lead to poor transformations to normality, as shown in Figure \ref{fig:sensitivity-to-outliers}. As indicated by @Raymaekers2021-kq, the general aim of power transformations should be to transform non-outlier data to normality, i.e. achieve *central normality*. To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$. Because $\lambda$ is updated during their procedure, simultaneous estimation of $\lambda$ and $x_0$ for shift-sensitive power transformations is not possible.
